A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-15 paper/draft.tex
test run data donations

Why should patronage occur post-election? There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office. In particular, a politician cannot offer job to a voter prior to holding office. Moreover, public sector jobs are coarse goods in the sense that they cannot be infinitely divided among voters. There are tight budget and administrative constraints that distinguishes a public sector jobs from other forms of resources available to politicians.

If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician. Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, constituencies can vary widely with respect to the resources that they bring in to the table. A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}. What is underemphasized is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.

@@

Additionally, these same workers tend to have a longer experience working in the formal sector prior to employment. Combined with the higher compensations, the employment evidence suggests that that party members who first enter the bureaucracy in the initial years of the mandate belong to a wealthier strata than their non-partisan counterparts.

@@

@@

In this section, I provide descriptive statistics that illustrate broader patterns in the distribution of party affiliated bureaucrats at the municipal level in Brazil. Figure \ref{fig:map_pooled} illustrates the subnational distribution of party affiliation in Brazil for the year of 2015. There are a few patterns that emerge.

>>>>

The timing of clientelistic transfers ie key. There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office. In particular, a politician cannot offer job to a voter prior to holding office. Moreover, public sector jobs are coarse goods in the sense that they cannot be infinitely divided among voters. There are tight budget and administrative constraints that distinguishes a public sector jobs from other forms of resources available to politicians.

If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician. Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, members of the constituency can vary widely with respect to the resources that they bring in to the table. A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}. What is underemphasized is the fact that politicians can use patronage to lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.

@@

Additionally, these same workers tend to have a longer experience working in the formal sector prior to employment. Combined with the higher compensations, the employment evidence suggests that that party members who first enter the bureaucracy in the initial years of the mandate belong to a wealthier strata than their non-partisan counterparts. They have experienced both longer bouts of employment in the formal sector, as well as higher compensations in their tenure. While it is clear that these hires are a reward for previous contributions, it is not clear whether a party member who is hired into the bureaucracy is more likely to donate to their party. \emph{How can I show that party members who are hired into the bureaucracy are more likely to donate?}

@@

In sum, employment evidence seems to suggest that party members who enter the bureaucracy in the first years of mandate come predominantly from

@@

In this section, I provide descriptive statistics that illustrate broader patterns in the distribution of party affiliated bureaucrats at the municipal level in Brazil. Figure \ref{fig:map_pooled} illustrates the subnational distribution of party affiliation in Brazil for the year of 2015. There are a few patterns that emerge. The first one is that there is
